---
title: "HIK Project Day -- Linnaruum: Day 1 (11.10.2017)"
subtitle: "Preliminary Research Report"
author: "Luis P. Prieto"
abstract: "This document describes the research questions and data gathered for the first session of the HIK project days with activities outside the classroom (dubbed 'Linnaruum'), that took place on the 11 October 2017. It also explores what kinds of questions can be answered with such data, so that the whole process can be improved for future iterations of the event."
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, error = F)

library(ggplot2)
library(data.table)
library(xlsx)
library(reshape2)
library(FactoMineR)
library(factoextra)
library(GGally)

source("../src/preprocessMultimodalTracker.R")

```

(brief intro: this is a collaboration between HIK, didactics, several days - repetitions, overall topic and characteristics)


# Research questions


# Methodology

## Data gathering


```{r preprocess}

####################### MULTIMODAL TRACKER DATA
# The data file(s)
mm_datafiles <- list.files("../data/raw/20171011/", pattern = "*multimodal-tracker*")
#filename <- "./multimodal-tracker-1507639281494.txt"

#TODO: Implement automated download from the raw data repository
# Assume we have all the raw data in ../data/raw
#overall <- data.frame()
acceleration <- data.frame()
beacons <- data.frame()
geolocation <- data.frame()

for(file in mm_datafiles){

   filedata <- preprocessMultimodalTracker(paste("../data/raw/20171011/",file,sep=""))
   
   group <- as.numeric(substr(file,2,2))
   # filedata$overall$group <- group
   # if(nrow(overall)==0) overall <- filedata$overall
   # else overall <- rbind(overall, filedata$overall)
   
   if(!is.null(filedata$acceleration)){
     filedata$acceleration$group <- group
     if(nrow(acceleration)==0) acceleration <- filedata$acceleration
     else acceleration <- rbind(acceleration, filedata$acceleration)
   }

   if(!is.null(filedata$beacons)){
     filedata$beacons$group <- group
     if(nrow(beacons)==0) beacons <- filedata$beacons
     else beacons <- rbind(beacons, filedata$beacons)
   }
   
   if(!is.null(filedata$geolocation)){
     filedata$geolocation$group <- group
     if(nrow(geolocation)==0) geolocation <- filedata$geolocation
     else geolocation <- rbind(geolocation, filedata$geolocation)
   }
}

# print("Geolocation data")
# print(xtabs(~group,data=geolocation))
# for(i in 1:6){
#   print(paste("Group",i))
#   times <- geolocation[geolocation$group==i,"timestamp"]
#   print(paste("We have samples from ",as.POSIXct(min(times)/1000, origin = "1970-01-01"),"to",as.POSIXct(max(times)/1000, origin = "1970-01-01")))
# }
# 
# print("Beacon data")
# print(xtabs(~group,data=beacons))
# for(i in 1:6){
#   print(paste("Group",i))
#   times <- beacons[beacons$group==i,"timestamp"]
#   print(paste("We have samples from ",as.POSIXct(min(times)/1000, origin = "1970-01-01"),"to",as.POSIXct(max(times)/1000, origin = "1970-01-01")))
# }
# 
# print("Acc data")
# print(xtabs(~group,data=acceleration))
# for(i in 1:6){
#   print(paste("Group",i))
#   times <- acceleration[acceleration$group==i,"timestamp"]
#   print(paste("We have samples from ",as.POSIXct(min(times)/1000, origin = "1970-01-01"),"to",as.POSIXct(max(times)/1000, origin = "1970-01-01")))
# }


########################## OBSERVER DATA
# The data file(s)
obs_datafile <- "../data/raw/20171011/Observer sheet - Linnaruum (Responses).xlsx"

obs_data <- read.xlsx(obs_datafile, sheetIndex = 1)
names(obs_data) <- c("timestamp", "group", "StudentA", "StudentB", "StudentC", "StudentD", 
                     "StudentE", "StudentF", "StudentG", "comment")

obs_data <- obs_data[,c(-8,-9)]

# Student view of the observations
student_obs <- melt(obs_data, id=1:2, measure=3:7)
student_obs$disengaged <- as.numeric(grepl(pattern = "disengaged", x = student_obs$value, fixed = TRUE))
student_obs$looking <- as.numeric(grepl(pattern = "Looking", x = student_obs$value, fixed = TRUE))
student_obs$talking <- as.numeric(grepl(pattern = "Talking", x = student_obs$value, fixed = TRUE))
student_obs$resources <- as.numeric(grepl(pattern = "technology", x = student_obs$value, fixed = TRUE))
student_obs$external <- as.numeric(grepl(pattern = "external", x = student_obs$value, fixed = TRUE))
names(student_obs)[[3]]<-"student"
student_obs <- student_obs[,c(1:3,5:9)]


# summary(student_obs)


# Group view of the observations
group_obs <- obs_data[,c("timestamp","group","comment")]
ag_dis <- aggregate(disengaged ~ timestamp+group, data=student_obs, FUN=sum)
ag_look <- aggregate(looking ~ timestamp+group, data=student_obs, FUN=sum)
ag_talk <- aggregate(talking ~ timestamp+group, data=student_obs, FUN=sum)
ag_res <- aggregate(resources ~ timestamp+group, data=student_obs, FUN=sum)
ag_ext <- aggregate(external ~ timestamp+group, data=student_obs, FUN=sum)

group_obs <- merge(group_obs, ag_dis)
group_obs <- merge(group_obs, ag_look)
group_obs <- merge(group_obs, ag_talk)
group_obs <- merge(group_obs, ag_res)
group_obs <- merge(group_obs, ag_ext)
group_obs <- group_obs[order(group_obs$group,group_obs$timestamp),]

group_obs$group <- as.factor(group_obs$group)
group_obs$comment <- as.character(group_obs$comment)
# summary(group_obs)
# names(group_obs)

# ag <- aggregate(cbind(disengaged,looking,talking,resources,external)~group, data=group_obs, FUN = mean)
# ag


############### STUDENT FEEDBACK
# Read and clean the individual activity feedback data
fb1_datafile <- "../data/raw/20171011/Avastame Kadriorgu - Individual activity feedback (Responses).xlsx"

fb1_data <- read.xlsx(fb1_datafile, sheetIndex = 1)
names(fb1_data) <- c("timestamp", "group", "student", "activity", "contribution", "difficulty", "preparedness", "satisfaction", "relevant")
levels(fb1_data$activity) <- c("I", "II", "III", "IV")
fb1_data$contribution <- as.numeric(gsub( " .*$", "", fb1_data$contribution ))
fb1_data$student.id <- paste(fb1_data$group,fb1_data$student)

# Read and clean the final feedback data
fb2_datafile <- "../data/raw/20171011/Avastame Kadriorgu - Final feedback (Responses).xlsx"

fb2_data <- read.xlsx(fb2_datafile, sheetIndex = 1)
names(fb2_data) <- c("timestamp", "group", "student", "strategies", "roles", "overall.satisfaction", "suggestions")
fb2_data$student.id <- paste(fb2_data$group,fb2_data$student)
```

## Data analysis

# The events of 11.10.2017

## What data is available?

## What kind of questions can we answer?


# How can we improve?

## The learning scenario


## The research process

Other data to be gathered

Other analyses to be run