---
title: "Mysterious Island -- Our data from multiple sources"
author: "Luis P. Prieto"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r helpers}
library(gsheet)
library(ggplot2)
library(wordcloud)
library(Hmisc)
library(tm)
library(SnowballC)
library(knitr)

# Some helper functions
source("../src/processObservationData.R")
source("../src/processFinalTaskData.R")
```

```{r datasources}

# We define and get the data from all our data sources, and merge it into master tables for convenience

# Observer sheets
url_obs1 = "https://docs.google.com/spreadsheets/d/1FSqm57ygxaIFa5V5oUCx7ht21ayuxMm7MO2IeczZ7ww/edit#gid=1838199304"
url_obs2 = "https://docs.google.com/spreadsheets/d/13Od5UuY5LLh2E5EPVD17IP6dO3x_xTpxyVYg9BQmIdE/edit#gid=140638193"
url_obs3 = "https://docs.google.com/spreadsheets/d/1hkkeSRYKKtlpxYAA4jEUddhkv-X5oj6gXzGPKobA5Wc/edit#gid=2089592913"
url_obs4 = NULL # No observation done! See quali observation in PDF file...

# Individual activity feedback
url_ind1 = "https://docs.google.com/spreadsheets/d/1NbsjX7vQqgfCKmTJQ-9Zs3zj2dU0CrnMBavXy9qZfqc/edit#gid=76827960"
url_ind2 = "https://docs.google.com/spreadsheets/d/13GZK27me6PYfo7AitvIvWVR_RmhhVP6ypGgc_VaqH2s/edit#gid=1439210366"
url_ind3 = "https://docs.google.com/spreadsheets/d/1o-JBf9MeReo8ob8LT80OfE3_ZQwFG5Xi9AJtRkxSRUc/edit#gid=1834453828"
url_ind4 = "https://docs.google.com/spreadsheets/d/11cDBX25XCJq89UdmWydXEfMvdHcA4xG-jhbSBgxEnQw/edit#gid=1341825476"

# Final day feedback
url_fin1 = "https://docs.google.com/spreadsheets/d/1v2E-FQMl4t2qJAAc3Vg8Qo6-7-W_Ajq5ja89QBPk4Fs/edit#gid=1280693886"
url_fin2 = "https://docs.google.com/spreadsheets/d/1JjXYqMGQmTKbAW9q_u7Rl9LWgglIOiVGVKU6iBTijl0/edit#gid=1575889659"
url_fin3 = "https://docs.google.com/spreadsheets/d/1CRUVSsgMcNdJ_kWf3T9aQfMvh3rSQe94fzZMoepOMvw/edit#gid=241912863"
url_fin4 = "https://docs.google.com/spreadsheets/d/1AyTGEA_Sb69Sf-m0HaiAczhr9jjjiFo57yJQVtHetIo/edit#gid=814282282"

# Graasp logs

# Other sources (e.g., multimodal tracker, student-generated artifacts from Graasp)?


```

The data we have:

* Observations: every 5min, both quantitative (how each student is engaged) and qualitative (what is happening generally with the group)
* Individual activity student feedback: satisfaction, difficulty, collaboration self-report after each of the activities
* Final whole-day student feedback: overall satisfaction, suggestions for activities, etc.
* Logs from activity in Graasp (when students connected, responded, accessed, etc.)
* Artifacts from Graasp (e.g., student responses)

```{r preprocess}

# Download, parse and merge
# Observations
rawobs1 <- as.data.frame(gsheet2tbl(url_obs1))
obs1 <- processObservationData(rawobs1, 
                               date = as.POSIXct(strptime("06-12-2017", "%d-%m-%Y")),
                               observercol = F)
rawobs2 <- as.data.frame(gsheet2tbl(url_obs2))
obs2 <- processObservationData(rawobs2, 
                               date = as.POSIXct(strptime("13-12-2017", "%d-%m-%Y")),
                               observercol = F)
rawobs3 <- as.data.frame(gsheet2tbl(url_obs3))
obs3 <- processObservationData(rawobs3, 
                               date = as.POSIXct(strptime("10-01-2018", "%d-%m-%Y")),
                               observercol = T,
                               namecols=c("timestamp","group","activity",
                                              "StudentA","StudentB","StudentC",
                                              "StudentD","StudentE","comment","observer"))
obs <- rbind(obs1,obs2)
obs <- merge(obs, obs3,all = T)
obs$timestamp <- as.POSIXct(strptime(obs$timestamp, "%d/%m/%Y %H:%M:%S"))
obs <- obs[ order(obs$timestamp), ]

# Individual activity fb
rawind1 <- as.data.frame(gsheet2tbl(url_ind1))
names(rawind1) <- c("timestamp", "group", "student", "activity", "contribution",
                    "difficulty", "preparedness", "satisfaction", "relevant", 
                    "collab.challenges", "skill.gap")
rawind2 <- as.data.frame(gsheet2tbl(url_ind2))
names(rawind2) <- c("timestamp", "group", "student", "activity", "contribution",
                    "difficulty", "preparedness", "satisfaction", "relevant", 
                    "collab.challenges", "skill.gap")
rawind3 <- as.data.frame(gsheet2tbl(url_ind3))
names(rawind3) <- c("timestamp", "group", "student", "activity", "contribution",
                    "difficulty", "preparedness", "satisfaction", "relevant", 
                    "collab.challenges", "skill.gap")
rawind4 <- as.data.frame(gsheet2tbl(url_ind4))
names(rawind4) <- c("timestamp", "group", "student", "activity", "contribution",
                    "difficulty", "preparedness", "satisfaction", "relevant", 
                    "collab.challenges", "skill.gap")
ind <- rbind(rawind1, rawind2, rawind3, rawind4)
ind$timestamp <- as.POSIXct(strptime(ind$timestamp, "%d/%m/%Y %H:%M:%S"))
ind <- ind[ order(ind$timestamp), ]
# Activity names are irregular, unify
ind$activity <- as.factor(ind$activity)
levels(ind$activity) <- c("Hõimuneem", "Hõimuneem", "Koordinaatide org", "Koordinaatide org",
                          "Ohukraater", "Ohukraater", "Pääsemise laht", "Pääsemise laht",
                          "Sissejuhatus", "Suitsusignaal")
# Dates
ind$date <- as.factor(as.character(as.Date(ind$timestamp)))


# Final activity fb
rawfin1 <- as.data.frame(gsheet2tbl(url_fin1))
names(rawfin1) <- c("timestamp", "group", "student", "strategies", "roles",
                    "overall.satisfaction", "other.activities", "ease.instructions",
                    "tech.problems", "other.problems")
rawfin1$timestamp <- as.POSIXct(strptime(rawfin1$timestamp, "%m/%d/%Y %H:%M:%S")) #This date is in different format!!??
rawfin1$timestamp <- format(rawfin1$timestamp, format = "%d/%m/%Y %H:%M%:%S")

rawfin2 <- as.data.frame(gsheet2tbl(url_fin2))
names(rawfin2) <- c("timestamp", "group", "student", "strategies", 
                    "tasks.student.A", "tasks.student.B", "tasks.student.C", 
                    "tasks.student.D", "tasks.student.E",
                    "roles", "overall.satisfaction", "other.activities", 
                    "ease.instructions", "tech.problems", "other.problems")
fin2 <- processFinalTaskData(rawfin2)

rawfin3 <- as.data.frame(gsheet2tbl(url_fin3))
names(rawfin3) <- c("timestamp", "group", "student", "strategies", 
                    "tasks.student.A", "tasks.student.B", "tasks.student.C", 
                    "tasks.student.D", "tasks.student.E",
                    "roles", "overall.satisfaction", "other.activities", 
                    "ease.instructions", "tech.problems", "other.problems")
fin3 <- processFinalTaskData(rawfin3)

rawfin4 <- as.data.frame(gsheet2tbl(url_fin4))
names(rawfin4) <- c("timestamp", "group", "student", "strategies", 
                    "tasks.student.A", "tasks.student.B", "tasks.student.C", 
                    "tasks.student.D", "tasks.student.E",
                    "roles", "overall.satisfaction", "other.activities", 
                    "ease.instructions", "tech.problems", "other.problems")
fin4 <- processFinalTaskData(rawfin4)

fin <- rbind(fin2, fin3, fin4)
fin <- merge(fin, rawfin1, all=T)
fin$timestamp <- as.POSIXct(strptime(fin$timestamp, "%d/%m/%Y %H:%M:%S"))
fin <- fin[ order(fin$timestamp), ]

```


## Overall view of the data

### Individual activity feedback


#### Perceived difficulty

Mean: `r mean(ind$difficulty)` (0-9 scale)

Median: `r median(ind$difficulty)` 

Std. deviation: `r sd(ind$difficulty)`


```{r ind-difficulty}

ggplot(ind, aes(x=difficulty))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived difficulty, overall")


ggplot(ind, aes(x=difficulty, fill=activity))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived difficulty, per activity")+
  facet_wrap(~activity)

ggplot(ind, aes(x=activity, y=difficulty, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  ggtitle("Distribution of student-perceived difficulty, per activity")

ggplot(ind, aes(x=activity, y=difficulty, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  facet_wrap(~date)+
  ggtitle("Distribution of student-perceived difficulty, per day")

```


#### Perceived preparedness

Mean: `r mean(ind$preparedness)` (0-9 scale)

Median: `r median(ind$preparedness)` 

Std. deviation: `r sd(ind$preparedness)`


```{r ind-preparedness}

ggplot(ind, aes(x=preparedness))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived preparedness, overall")


ggplot(ind, aes(x=preparedness, fill=activity))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived preparedness, per activity")+
  facet_wrap(~activity)

ggplot(ind, aes(x=activity, y=preparedness, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  ggtitle("Distribution of student-perceived preparedness, per activity")


ggplot(ind, aes(x=activity, y=preparedness, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  facet_wrap(~date)+
  ggtitle("Distribution of student-perceived preparedness, per day")

```


#### Perceived satisfaction

Mean: `r mean(ind$satisfaction)` (0-9 scale)

Median: `r median(ind$satisfaction)` 

Std. deviation: `r sd(ind$satisfaction)`


```{r ind-satisfaction}

ggplot(ind, aes(x=satisfaction))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived satisfaction, overall")


ggplot(ind, aes(x=satisfaction, fill=activity))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived satisfaction, per activity")+
  facet_wrap(~activity)

ggplot(ind, aes(x=activity, y=satisfaction, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  ggtitle("Distribution of student-perceived satisfaction, per activity")

ggplot(ind, aes(x=activity, y=satisfaction, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  facet_wrap(~date)+
  ggtitle("Distribution of student-perceived satisfaction, per day")

```



#### Perceived relevant

Mean: `r mean(ind$relevant)` (0-9 scale)

Median: `r median(ind$relevant)` 

Std. deviation: `r sd(ind$relevant)`


```{r ind-relevant}

ggplot(ind, aes(x=relevant))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived relevant, overall")


ggplot(ind, aes(x=relevant, fill=activity))+geom_histogram(bins=10)+theme_minimal()+
  scale_x_continuous(breaks=c(0:9))+
  ggtitle("Student-perceived relevant, per activity")+
  facet_wrap(~activity)

ggplot(ind, aes(x=activity, y=relevant, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  ggtitle("Distribution of student-perceived relevant, per activity")

ggplot(ind, aes(x=activity, y=relevant, fill=activity))+
  geom_boxplot()+theme_minimal()+
  scale_y_continuous(breaks=c(0:9))+
  facet_wrap(~date)+
  ggtitle("Distribution of student-perceived relevant, per day")

```

#### Collaboration challenges (quali)

```{r ind-coll}
clean_vector <- function(v){
    newv <- v[v!=""]
    newv2 <- newv[nchar(newv)>1]
    newv2
}

coll <- ind$collab.challenges
d <- unlist(string.break.line(paste(coll,sep="\n")))
d <- clean_vector(d)
if(length(d)>0){
  corpus <- Corpus(VectorSource(d))
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, tolower)
  #corpus <- tm_map(corpus, removeWords, stopwords("english"))
  corpus <- tm_map(corpus, stemDocument)
  #corpus <- tm_map(corpus, PlainTextDocument)
  wordcloud(corpus, max.words=50, random.order=F)
}

yeses <- length(d[grepl(x=tolower(d), pattern="yes", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="jah", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="Да", fixed = T)])

noses <- length(d[grepl(x=tolower(d), pattern="no", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="ei", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="нет", fixed = T)])
  
print(paste("Yes-like responses (faced some kind of collaboration challenge): ",yeses,"out of",nrow(ind),"responses"))
print(paste("No-like responses (did not have collaboration problems): ",noses,"out of",nrow(ind),"responses"))

dprint <- d[nchar(d)>5]
kable(dprint, caption = "(longer) Responses")
```




#### Perceived missing skills (skill gap)


```{r ind-gap}

gap <- ind$skill.gap
d <- unlist(string.break.line(paste(gap,sep="\n")))
d <- clean_vector(d)
if(length(d)>0){
  corpus <- Corpus(VectorSource(d))
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, tolower)
  #corpus <- tm_map(corpus, removeWords, stopwords("english"))
  corpus <- tm_map(corpus, stemDocument)
  #corpus <- tm_map(corpus, PlainTextDocument)
  wordcloud(corpus, max.words=50, random.order=F)
}

yeses <- length(d[grepl(x=tolower(d), pattern="yes", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="jah", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="Да", fixed = T)])

noses <- length(d[grepl(x=tolower(d), pattern="no", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="ei", fixed = T)])+
  length(d[grepl(x=tolower(d), pattern="нет", fixed = T)])
  
print(paste("Yes-like responses (said they needed more skills): ",yeses,"out of",nrow(ind),"responses"))
print(paste("No-like responses (did not need more skills): ",noses,"out of",nrow(ind),"responses"))

dprint <- d[nchar(d)>5]
kable(dprint, caption = "(longer) Responses")
```

### Final whole-day feedback

### Observations



