---
title: "Inter Rater Reliability in project day (10.01.2018)"
author: "Luis P. Prieto"
date: "11 January 2018"
abstract: The data from the observations done on this week's project day indicates there was a MODERATE agreement between observers. This is probably not too bad for rather untrained observers, but may be worth looking into more carefully if our results rely heavily on this particular categorical observations.
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message = F)

library(irr)
library(ggplot2)
library(FactoMineR)
library(factoextra)
library(gsheet)
library(reshape2)
library(dplyr)
library(lubridate)
library(knitr)

```

## Inter-rater reliability (IRR) of observational data

... calculated from the data in the project day on 10.01.2018, in which four observers looked a two groups each. I calculated two typical metrics of IRR:

* Cohen's unweighted Kappa
* Krippendorff's Alpha

```{r pressure, echo=FALSE}

processObservationData <- function(data, 
                                   nrlevels=6, 
                                   date=as.POSIXct(strptime("10-01-2018", "%d-%m-%Y")),
                                   project="Isle", 
                                   activitycol=T,
                                   observercol=T){
  
  obs_data <- data
  
  #Clean up empty columns
  # for(i in ncol(obs_data):1){
  #   if(sum(complete.cases(obs_data[i]))==0 || sum(obs_data[i]!="")==0 || sum(obs_data[i]!="")==1) obs_data <- obs_data[,-i]
  # }
  
  # Cleanup invalid data, before 10am like "2017-10-11 09:44:49 GMT"
  #obs_data <- obs_data[obs_data$timestamp>as.POSIXct("2017-10-11 10:00:00 GMT", origin = "1970-01-01", tz = "GMT"),]
  
  # Student view of the observations
  if(observercol){
    if(!activitycol){
      student_obs <- melt(obs_data, id=c(1:2,ncol(obs_data)), measure=3:(ncol(obs_data)-2))
      names(student_obs)[[4]]<-"student"
    }else{
      student_obs <- melt(obs_data, id=c(1:3,ncol(obs_data)), measure=4:(ncol(obs_data)-2))
      names(student_obs)[[5]]<-"student"
    }
  }else{
    if(!activitycol){
      student_obs <- melt(obs_data, id=1:2, measure=3:(ncol(obs_data)-1))
      names(student_obs)[[3]]<-"student"
    }else{
      student_obs <- melt(obs_data, id=1:3, measure=4:(ncol(obs_data)-1))
      names(student_obs)[[4]]<-"student"
    }
  }
  student_obs$disengaged <- as.numeric(grepl(pattern = "disengaged", x = student_obs$value, fixed = TRUE))
  student_obs$looking <- as.numeric(grepl(pattern = "Looking", x = student_obs$value, fixed = TRUE))
  student_obs$talking <- as.numeric(grepl(pattern = "Talking", x = student_obs$value, fixed = TRUE))
  student_obs$technology <- as.numeric(grepl(pattern = "technology", x = student_obs$value, fixed = TRUE))
  student_obs$resources <- as.numeric(grepl(pattern = "resources", x = student_obs$value, fixed = TRUE))
  student_obs$external <- as.numeric(grepl(pattern = "external", x = student_obs$value, fixed = TRUE))
  student_obs$student.id <- paste(student_obs$group,student_obs$student)
  
  if(observercol){
    if(!activitycol){
      student_obs <- student_obs[,c(1:4,6:ncol(student_obs))]
    }else{
      student_obs <- student_obs[,c(1:5,7:ncol(student_obs))]
    }
  }else{
    if(!activitycol){
      student_obs <- student_obs[,c(1:3,5:ncol(student_obs))]
    }else{
      student_obs <- student_obs[,c(1:4,6:ncol(student_obs))]
    }
  }
  # summary(student_obs)
  
  student_obs$project <- project
  student_obs$date <- date
  
  student_obs$global.id <- paste(student_obs$project,student_obs$date,student_obs$student.id)
  
  student_obs  
}

getTimeofday <- function(data){
  data$time <- as.POSIXct(strptime(data$timestamp, "%d/%m/%Y %H:%M:%S"))
  data$timeofday <- as.POSIXct(strptime(paste(hour(data$time),
                                              minute(data$time),
                                              second(data$time), sep=":"),
                                        "%H:%M:%S"))
  data$timeofday
}



url_obs <- "https://docs.google.com/spreadsheets/d/1hkkeSRYKKtlpxYAA4jEUddhkv-X5oj6gXzGPKobA5Wc/edit"

data <- as.data.frame(gsheet2tbl(url_obs))
names(data) <- c("timestamp","group","activity",
                 "StudentA","StudentB","StudentC",
                 "StudentD","StudentE","comment","observer")

df <- processObservationData(data)
df$timeofday <- getTimeofday(df)

#Catch: always put the bigger data on data2!
matchObs <- function(data1, data2, col=3){
  data <- data1
  names(data)[col] <- paste(names(data)[col],".A",sep="")
  nameb <- paste(names(data2)[col],".B",sep="")
  data[,nameb] <- NA
  data[,"timeofday.b"] <- NA
  data[,"difftime"] <- NA
  for(i in 1:nrow(data)){
    d2 <- data2[data2$student.id==data[i,"student.id"],]
    diffs <- abs(d2[,1]-data[i,1])
    idx <- which.min(diffs)
    data[i,nameb] <- data2[row.names(d2[idx,]),col]
    data[i,"timeofday.b"] <- as.POSIXct(data2[row.names(d2[idx,]),"timeofday"])
    data[i,"difftime"] <- min(diffs)
  }
  
  data
}

# IRR calculations
results <- data.frame()
for(i in 6:11){ #For each of the variables disengaged, looking, etc
  #print(names(df)[i])
  results[nrow(results)+1,"var"] <- names(df)[i]
  ob1a <- df %>% filter(observer=="1-A") %>% select(timeofday,student.id,i)
  ob1b <- df %>% filter(observer=="1-B") %>% select(timeofday,student.id,i)
  gr1 <- matchObs(ob1b,ob1a)
  results[nrow(results),"group"] <- "Group1"
  # Cohen's Kappa
  k1 <- kappa2(gr1[,c(3,4)], "unweighted")
  #print(kappa2(gr1[,c(3,4)], "unweighted"))
  results[nrow(results),"kappa"] <- k1$value
  results[nrow(results),"kappa.p"] <- k1$p.value
  # Krippendorf's alpha
  a1 <- kripp.alpha(t(as.matrix(gr1[,c(3,4)])), method="nominal")
  #print(kripp.alpha(t(as.matrix(gr1[,c(3,4)])), method="nominal"))
  results[nrow(results),"alpha"] <- a1$value

  results[nrow(results)+1,"var"] <- names(df)[i]
  ob2a <- df %>% filter(observer=="2-A") %>% select(timeofday,student.id,i)
  ob2b <- df %>% filter(observer=="2-B") %>% select(timeofday,student.id,i)
  gr2 <- matchObs(ob2a,ob2b)
  results[nrow(results),"group"] <- "Group2"
  # Cohen's Kappa
  k2 <- kappa2(gr2[,c(3,4)], "unweighted")
  #print(kappa2(gr2[,c(3,4)], "unweighted"))
  results[nrow(results),"kappa"] <- k2$value
  results[nrow(results),"kappa.p"] <- k2$p.value
  # Krippendorf's alpha
  a2 <- kripp.alpha(t(as.matrix(gr2[,c(3,4)])), method="nominal")
  #print(kripp.alpha(t(as.matrix(gr2[,c(3,4)])), method="nominal"))
  results[nrow(results),"alpha"] <- a2$value
  
  #irrdata <- rbind(gr1,gr2)
}

names(results)[1] <- "behavior"
names(results)[4] <- "p-value (prob. observations were by chance)"


kable(results)

# All results
ggplot(results, aes(x="ALL", y=kappa))+geom_boxplot()+theme_minimal()
ggplot(results, aes(x="ALL", y=alpha))+geom_boxplot()+theme_minimal()

# By kind of observation
ggplot(results, aes(x=behavior, y=kappa))+geom_boxplot()+theme_minimal()
ggplot(results, aes(x=behavior, y=alpha))+geom_boxplot()+theme_minimal()


# By group (set of observers)
ggplot(results, aes(x=group, y=kappa))+geom_boxplot()+theme_minimal()
ggplot(results, aes(x=group, y=alpha))+geom_boxplot()+theme_minimal()



```

## Conclusions

As we can see from the result table and graphs, the inter-rater reliability values are moderate (Kappa and Alpha around 0.4--0.6), with one strange case of worse-than-chance agreement in Group 1 for the "disengaged" behavior.

In general (avg. Kappa=`r mean(results$kappa)`, avg. Alpha=`r mean(results$alpha)`), this is probably not too bad for rather untrained observers, but may be worth looking into more carefully if our results rely heavily on this particular categorical observations.
