---
title: "Processing Observation Data from Learning Experiments"
author: "Lewis Marsh and Luis P. Prieto"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gsheet)
library(knitr)

source("../src/processObservationData.R")
source("../src/mca.R")

raw_data <- as.data.frame(gsheet2tbl("https://docs.google.com/spreadsheets/d/11BPHcSlqwozx3ffOQhQ5i5O8TRVa8xKoaw2uiny8k5A/edit"))
```

This R Markdown report aims to summarise the analyses done on the observation data of several sessions of innovative learning activities, from which multimodal data has been gathered

# Observation data

## Preprocessing

The preprocessing is taken care by the R script in `src/processObservationData.R`. It consists of two main functions, `processObservationData` (to process the data of one session) and `processAllObservationData` (for multiple sessions).

The first of these two functions,
```
data <- processObservationData(raw_data,
                       date=as.POSIXct(strptime("10-01-2018", "%d-%m-%Y")),
                       project="Isle", 
                       activitycol=F,
                       observercol=F,
                       namecols=c("timestamp","group",
                            "StudentA","StudentB","StudentC","StudentD"))
```
takes in a dataframe with the raw observation data, together with some other parameters about the structure of the raw dataset and expected columns in the processed dataset, and returns a clean data frame with the observations for each individual.

In the orginal data sheets, each entry lists the observation for a whole group of students (between 3 and 6
usually) and the observations are listed as a piece of (predefined) text, such as "Talking with their group peers
to solve the task" or "Totally disengaged".

In the output this is broken up into individual entries for each student, for which a student variable is
introduced. Furthermore, each possible observation is turned into a boolean variable for each of these entries,
which makes it easier to conduct further statistical analysis of the dataset. Also, a column
with additional notes from the original dataframe is dropped in the output, as it is difficult to make use of
in a statistical test, and a global ID is created for each student, using the date, group, student number in the
group and the project name. This will be of use when merging several datasets with the second function.

As students undertook several tastks in some of the experiments/projects, the input variable "activitycol" indicates whether the function has to look for an extra activity column in the input data. This will be taken over
one-to-one in the output or will be filled in with "Standard" if there was only one activity.
Similarly, in some projects there were more than one observer placed onto each group to enable to test the
reliability of the observers and the data they produce. The variable "observercol" tells the function whether
there is a separate observer column to take over from the input data. Otherwise "1-A" is filled in for each
observer.

The second function, `processAllObservationData` takes in a vector of URLs in which the GoogleDocs
containing the experimental data are stored in. It reads the data from the online sheets and places
them in a dataframe. By examining the names of the columns in the dataframe, it detects whether there
are activity or observation columns. By inspecting the first time stamp, it determines the date of the experiment.
It then cleans the data using the first function and adds it to the end of a larger dataframe which
eventually gets returned after the function has iterated through all URLs.

### Example of Output
To exhibit what the functions do, there is a brief example. We are given a data frame taken from
one of the GoogleDocs:
```{r first, include=FALSE}
raw_data <- as.data.frame(gsheet2tbl("https://docs.google.com/spreadsheets/d/11BPHcSlqwozx3ffOQhQ5i5O8TRVa8xKoaw2uiny8k5A/edit"))
```
```{r second}
raw_data[1,1:2]
raw_data[1,3]
```

These are only two snippets of the data, as anything else would not fit onto this page.
We see that the first two columns give a timestamp and the group. We furthermore see the oberservation for
student A. There are 3 other students in the group with individual columns as well as
a column for further comments.

By applying the `processObservationData` function, we then see how the student variable is created
and the observations are turned from text into boolean values. Furthermore, the activity and
observer column have been created which makes it easier to merge and compare this data with other datasets.
```{r third}
processed_data <- processAllObservationData()
summary(processed_data)
str(processed_data)
```

### Spot checking the data

Some tables to see what kind of data we have

```{r fourth}

table(processed_data$date, processed_data$student)

```


## MCA

The `mca.R` contains several functions to create and interpret an MCA conducted on the
idicator variables created above.

Firstly, `mca_analysis(data, plots = T)` returns an mca element of the `FactoMineR` package.
It has the options to create several plots, such as a bar plot of the significance of each newly created
MCA dimension, the variance of the old variables in the first two dimensions and the distribution of
each variable from the data set across the first two dimensions. An example of the last instance:

```{r fifth, include = T}
  mca <- mca_analysis(processed_data, plots = F)
  fviz_ellipses(mca, "disengaged", geom = "point")
```

A further function is `add_mca_dimensions(dataframe, mca = NULL, dimensions = 3)`. It merges
a chosen amount of dimension generated by the MCA into the original dataframe. One can choose
whether one wants to generated the MCA object separately using the first functions
(e.g. to inspect the plots first) or let the function conduct the MCA itself (if no mca is passed).
For example:

```{r sixth, include = T}
  processed_data <- add_mca_dimensions(processed_data, mca = mca, dimensions = 3)
```

There are several functions that visualise the distribution of our given data points. All of these take
in a dataframe to which the MCA dimensions have been added.

For instance `create_dimension_histograms(dataframe, dimensions = 3)` creates a series of histograms
that exhibit the distribution of data points on a desired amount of variables (by default the first 3).
Similarly thereto, `reate_date_histograms(dataframe, dimension = 1)` creates histograms that show
the distrbution of data points on a chosen dimension with each plot representing one day.

The information conatined in the histograms distingushing dates can also be shown in one single plot,
using the `create_violin_plot` function. For our specific data set, this will looke like:

```{r seventh}
  create_violin_plot(processed_data, dimension = 1)
```

Finally, the function `aggregate_by_groups(dataframe)` allows us to aggregate the dataframe
on groups and timestamps, summing up the MCA dimension.

```{r eighth}
  processed_data <- aggregate_by_groups(processed_data)
  summary(processed_data)
```


## Topic modeling of observer comments

... using Latent Dirichlet Allocation (LDA) (TBD by Luis)
