---
title: "Analyzing project days' engagement v2"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gsheet)
library(knitr)
library(dplyr)
library(magrittr)
library("PerformanceAnalytics")
library(Hmisc)

source("../src/processObservationData.R")
source("../src/mca.R")
source("../src/hmm.R")
source("../src/lda.R")
source("../src/auto_encoder.R")
source("../src/questionnaires.R")
```

We import the observation and questionnaire data...

```{r, import, warning=FALSE, message=FALSE, echo=FALSE}

obsdata <- processAllObservationData()
actdata <- load_all_intermediate_questionnaires()
findata <- load_all_final_questionnaires()

```

Initial explorations of observation data:

```{r}

obsdata %>% dplyr::select(date, project, disengaged:external) %>% 
  aggregate(.~date+project,data=., mean) %>%
  dplyr::mutate(date=as.factor(date)) %>%
  melt(id.vars=c("date","project")) %>%
  ggplot(aes(date, value)) + 
  geom_bar(aes(fill=variable), position = "dodge", stat = "identity") + theme_minimal()


obsdata %>% dplyr::select(project, disengaged:external) %>% aggregate(.~project,data=., mean) %>%
  melt(id.vars="project") %>%
  ggplot(aes(project, value)) + 
  geom_bar(aes(fill=variable), position = "dodge", stat = "identity") + theme_minimal()

```

*One thing to note: on the 2017-10-11 event, the observation of technology and resources was conflated into a single observer tick -- the current preprocessing assumes that the tick means "engaged with technology AND resources" (and the absence of tick, not engaged with either technology or resources)*

We can see that different dates have quite different engagement profiles, and the same can be said for the aggregation of every kind of project day: e.g., in Keha there was more use of technology and resources, while in Linnaruum there was more talking. However, this is not so easy to interpret, especially for a teacher concerned by the question: "are my students engaged?"

A potential way of answering this with MCA analysis:

```{r}

mca <- mca_analysis(obsdata, plots = T)
#  fviz_ellipses(mca, "disengaged", geom = "point")
obsdata <- add_mca_dimensions(obsdata, mca = mca, dimensions = 3)

obsdata <- obsdata %>% mutate(MCA.Active.Eng = -MCAdim1) %>%
   mutate(MCA.Passive.Eng = -MCAdim2) %>%
   mutate(MCA.Conf.Help = MCAdim3)

```

* MCADim1 explains most (25%) of the variance, and it seems very correlated with disengagement behavior (and somewhat with more passive forms of engagement like looking), and viceversa: negative values are related with the other, more active kinds of engagement (external, resources, technology, talking). We can interpret --MCADim1 as **"Active engagement"**. 
* --MCADim2 could be related with some kind of **"Passive engagement"** (introspection?), since it is quite correlated with looking and talking (but not use of external, techology, or disengagement). 
* MCADim3 is related with external and looking, maybe in the sense of **"confusion and help-seeking"**? (but this is maybe a stretch)


With these interpretations, we can now get the profile of each day, each project, and each student or group:

```{r}
obsdata %>% dplyr::select(date, project, MCA.Active.Eng:MCA.Conf.Help) %>% 
  aggregate(.~date+project,data=., mean) %>%
  dplyr::mutate(date=as.factor(date)) %>%
  melt(id.vars=c("date","project")) %>%
  ggplot(aes(date, value)) + 
  geom_bar(aes(fill=variable), position = "dodge", stat = "identity") + theme_minimal()


obsdata %>% dplyr::select(project, MCA.Active.Eng:MCA.Conf.Help) %>% aggregate(.~project,data=., mean) %>%
  melt(id.vars="project") %>%
  ggplot(aes(project, value)) + 
  geom_bar(aes(fill=variable), position = "dodge", stat = "identity") + theme_minimal()

obsdata %>% dplyr::select(global.id, MCA.Active.Eng:MCA.Conf.Help) %>% aggregate(.~global.id,data=., mean) %>%
  melt(id.vars="global.id") %>%
  filter(global.id=="2017-08-11 Group 1 Student C" | global.id=="2017-08-11 Group 3 Student A") %>%
  ggplot(aes(global.id, value)) +
  geom_bar(aes(fill=variable), position = "dodge", stat = "identity") + theme_minimal()

obsdata %>% mutate(global.group.id=paste(date,group)) %>%
  dplyr::select(global.group.id, MCA.Active.Eng:MCA.Conf.Help) %>% aggregate(.~global.group.id,data=., mean) %>%
  melt(id.vars="global.group.id") %>%
  filter(global.group.id=="2017-08-11 Group 1" | global.group.id=="2017-08-11 Group 3") %>%
  ggplot(aes(global.group.id, value)) +
  geom_bar(aes(fill=variable), position = "dodge", stat = "identity") + theme_minimal()

```

... and we can also see the evolution of these variables over time, globally, per student and per group

```{r}

obsdata %>% dplyr::mutate(time=as.POSIXct(timestamp, tz="Europe/Tallinn", "%d/%m/%Y %H:%M:%S")) %>%
  dplyr::mutate(date=paste0(date," (",project,")")) %>%
  dplyr::select(time, date, project, MCA.Active.Eng:MCA.Conf.Help) %>% 
  melt(id.vars=c("time", "date","project")) %>%
  ggplot(aes(time, value, color=variable)) + 
  #geom_line(aes(color=variable)) + geom_point(aes(shape=project, color=variable)) + 
  geom_smooth(aes(color=variable), se = F, span=0.2) +
  facet_wrap(~date, scales = "free_x") + theme_minimal()

obsdata %>% dplyr::mutate(time=as.POSIXct(timestamp, tz="Europe/Tallinn", "%d/%m/%Y %H:%M:%S")) %>%
  dplyr::select(global.id, time, MCA.Active.Eng:MCA.Conf.Help) %>% 
  melt(id.vars=c("global.id","time")) %>%
  filter(global.id=="2017-08-11 Group 1 Student C" | global.id=="2017-08-11 Group 3 Student A") %>%
  ggplot(aes(time, value, color=variable)) + 
  #geom_line(aes(color=variable)) + geom_point(aes(shape=project, color=variable)) + 
  geom_smooth(aes(color=variable), se = F, span=0.2) + geom_rug(sides="b") +
  facet_wrap(~global.id, scales = "free_x") + theme_minimal()

obsdata %>% mutate(global.group.id=paste(date,group)) %>%
  dplyr::mutate(time=as.POSIXct(timestamp, tz="Europe/Tallinn", "%d/%m/%Y %H:%M:%S")) %>%
  dplyr::select(global.group.id, time, MCA.Active.Eng:MCA.Conf.Help) %>% 
  aggregate(.~time+global.group.id,data=., mean) %>%
  melt(id.vars=c("global.group.id","time")) %>%
  filter(global.group.id=="2017-08-11 Group 1" | global.group.id=="2017-08-11 Group 3") %>%
  ggplot(aes(time, value, color=variable)) + 
  #geom_line(aes(color=variable)) + geom_point(aes(shape=project, color=variable)) + 
  geom_smooth(aes(color=variable), se = F, span=0.2) + geom_rug(sides="b") +
  facet_wrap(~global.group.id, scales = "free_x") + theme_minimal()


```

If we now compare these MCA dimensions of engagement with the responses to the final questionnaire (difficulty --actually, ease; and satisfaction):

```{r, eval=F}

obtotal <- obsdata %>% dplyr::select(global.id, disengaged:external, MCA.Active.Eng:MCA.Conf.Help) %>% aggregate(.~global.id,data=., mean)

merged <- merge(obtotal,findata, by="global.id") %>% dplyr::select(global.id, disengaged:external, MCA.Active.Eng:MCA.Conf.Help, satisfaction, difficulty)

merged %>% dplyr::select(-global.id) %>% chart.Correlation(histogram=TRUE, pch=19)
```

The only interesting correlations are ease-confusion.help (confirms a bit the interpretation) and ease-looking (if it is easy, they don't look, just act?)

Do these dimensions (or the autoencoder) take away some of the observer noise (in IRR)? Apparently not

## Adding the HMM and auto-encoder variables

```{r}
#create_hmm_states_global(obsdata, visualise = TRUE, max_states = 6)
obsdata <- insert_hmm(obsdata)

training_and_test <- obsdata %>% filter(row_number() <= nrow(obsdata)*0.75)
validation <- obsdata %>% filter(row_number() > nrow(obsdata)*0.75)

model <- build_autoencoder(training_and_test)

predictions <- get_predictions(validation, model)

summary(predictions$AEerr)
obsdata <- insert_ae_units(obsdata, model)

```

## Comparing with activity-level questionnaires

```{r}
# See https://rstudio-pubs-static.s3.amazonaws.com/240657_5157ff98e8204c358b2118fa69162e18.html for multiple correlations tests


matched_data <- match_with_data(obsdata, iq=actdata, fq=findata)
agg_data <- merge_and_aggregate(matched_data, actdata, findata)

agg_data %>% dplyr::select(disengaged:external, MCAdim1:AEdim3, contribution:collaboration,f_satisfaction,ease) %>% as.matrix() %>% rcorr(type = c("pearson","spearman")) %>% kable()

```