---
title: "Project Days (11.10.2017) - Exploratory Data Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)

library(ggplot2)
library(data.table)
library(xlsx)
library(reshape2)
library(FactoMineR)
library(factoextra)
library(GGally)

source("../src/preprocessMultimodalTracker.R")
```

# Preprocessing

## Read/preprocess the multimodal tracker data


```{r multi-preprocess}
# The data file(s)
mm_datafiles <- list.files("../data/raw/20171011/", pattern = "*multimodal-tracker*")
#filename <- "./multimodal-tracker-1507639281494.txt"

#TODO: Implement automated download from the raw data repository
# Assume we have all the raw data in ../data/raw
#overall <- data.frame()
acceleration <- data.frame()
beacons <- data.frame()
geolocation <- data.frame()

for(file in mm_datafiles){

   filedata <- preprocessMultimodalTracker(paste("../data/raw/20171011/",file,sep=""))
   
   group <- as.numeric(substr(file,2,2))
   # filedata$overall$group <- group
   # if(nrow(overall)==0) overall <- filedata$overall
   # else overall <- rbind(overall, filedata$overall)
   
   if(!is.null(filedata$acceleration)){
     filedata$acceleration$group <- group
     if(nrow(acceleration)==0) acceleration <- filedata$acceleration
     else acceleration <- rbind(acceleration, filedata$acceleration)
   }

   if(!is.null(filedata$beacons)){
     filedata$beacons$group <- group
     if(nrow(beacons)==0) beacons <- filedata$beacons
     else beacons <- rbind(beacons, filedata$beacons)
   }
   
   if(!is.null(filedata$geolocation)){
     filedata$geolocation$group <- group
     if(nrow(geolocation)==0) geolocation <- filedata$geolocation
     else geolocation <- rbind(geolocation, filedata$geolocation)
   }
}

print("Geolocation data")
print(xtabs(~group,data=geolocation))
for(i in 1:6){
  print(paste("Group",i))
  times <- geolocation[geolocation$group==i,"timestamp"]
  print(paste("We have samples from ",as.POSIXct(min(times)/1000, origin = "1970-01-01"),"to",as.POSIXct(max(times)/1000, origin = "1970-01-01")))
}

print("Beacon data")
print(xtabs(~group,data=beacons))
for(i in 1:6){
  print(paste("Group",i))
  times <- beacons[beacons$group==i,"timestamp"]
  print(paste("We have samples from ",as.POSIXct(min(times)/1000, origin = "1970-01-01"),"to",as.POSIXct(max(times)/1000, origin = "1970-01-01")))
}

print("Acc data")
print(xtabs(~group,data=acceleration))
for(i in 1:6){
  print(paste("Group",i))
  times <- acceleration[acceleration$group==i,"timestamp"]
  print(paste("We have samples from ",as.POSIXct(min(times)/1000, origin = "1970-01-01"),"to",as.POSIXct(max(times)/1000, origin = "1970-01-01")))
}



```

## Read the observer data


```{r obs-preprocess}
# The data file(s)
obs_datafile <- "../data/raw/20171011/Observer sheet - Linnaruum (Responses).xlsx"

obs_data <- read.xlsx(obs_datafile, sheetIndex = 1)
names(obs_data) <- c("timestamp", "group", "StudentA", "StudentB", "StudentC", "StudentD", 
                     "StudentE", "StudentF", "StudentG", "comment")

obs_data <- obs_data[,c(-8,-9)]

# Student view of the observations
student_obs <- melt(obs_data, id=1:2, measure=3:7)
student_obs$disengaged <- as.numeric(grepl(pattern = "disengaged", x = student_obs$value, fixed = TRUE))
student_obs$looking <- as.numeric(grepl(pattern = "Looking", x = student_obs$value, fixed = TRUE))
student_obs$talking <- as.numeric(grepl(pattern = "Talking", x = student_obs$value, fixed = TRUE))
student_obs$resources <- as.numeric(grepl(pattern = "technology", x = student_obs$value, fixed = TRUE))
student_obs$external <- as.numeric(grepl(pattern = "external", x = student_obs$value, fixed = TRUE))
names(student_obs)[[3]]<-"student"
student_obs <- student_obs[,c(1:3,5:9)]

summary(student_obs)


# Group view of the observations
group_obs <- obs_data[,c("timestamp","group","comment")]
ag_dis <- aggregate(disengaged ~ timestamp+group, data=student_obs, FUN=sum)
ag_look <- aggregate(looking ~ timestamp+group, data=student_obs, FUN=sum)
ag_talk <- aggregate(talking ~ timestamp+group, data=student_obs, FUN=sum)
ag_res <- aggregate(resources ~ timestamp+group, data=student_obs, FUN=sum)
ag_ext <- aggregate(external ~ timestamp+group, data=student_obs, FUN=sum)

group_obs <- merge(group_obs, ag_dis)
group_obs <- merge(group_obs, ag_look)
group_obs <- merge(group_obs, ag_talk)
group_obs <- merge(group_obs, ag_res)
group_obs <- merge(group_obs, ag_ext)
group_obs <- group_obs[order(group_obs$group,group_obs$timestamp),]

group_obs$group <- as.factor(group_obs$group)
group_obs$comment <- as.character(group_obs$comment)
summary(group_obs)
names(group_obs)

ag <- aggregate(cbind(disengaged,looking,talking,resources,external)~group, data=group_obs, FUN = mean)
ag
```


## Read the student feedback data

```{r fb-preprocess}
# Read and clean the individual activity feedback data
fb1_datafile <- "../data/raw/20171011/Avastame Kadriorgu - Individual activity feedback (Responses).xlsx"

fb1_data <- read.xlsx(fb1_datafile, sheetIndex = 1)
names(fb1_data) <- c("timestamp", "group", "student", "activity", "contribution", "difficulty", "preparedness", "satisfaction", "relevant")
levels(fb1_data$activity) <- c("I", "II", "III", "IV")
fb1_data$contribution <- as.numeric(gsub( " .*$", "", fb1_data$contribution ))
fb1_data$student.id <- paste(fb1_data$group,fb1_data$student)

# Read and clean the final feedback data
fb2_datafile <- "../data/raw/20171011/Avastame Kadriorgu - Final feedback (Responses).xlsx"

fb2_data <- read.xlsx(fb2_datafile, sheetIndex = 1)
names(fb2_data) <- c("timestamp", "group", "student", "strategies", "roles", "overall.satisfaction", "suggestions")
fb2_data$student.id <- paste(fb2_data$group,fb2_data$student)
```


# Exploring the data

## Draw some plots - student feedback

```{r fb-plots1}
# How many responses we got?
table(fb1_data$group,fb1_data$activity)
table(fb1_data$student.id,fb1_data$activity)

plotdata <- fb1_data[,2:10]
melted <- melt(plotdata, id.vars = c(1:3,9), measure.vars = 4:8)

melted_cont <- melted[melted$variable=="contribution",]
melted_rest <- melted[melted$variable!="contribution",]

# Overall distribution of contributions, satisfaction, difficulty, preparedness
ggplot(melted_cont, aes(x=value))+geom_density()+theme_minimal()+xlim(0,100)
ggplot(melted_rest, aes(x=value))+geom_density()+theme_minimal()+ facet_grid(~ variable)+xlim(0,9)

# difficulty, preparedness and satisfaction per activity
ggplot(melted_rest, aes(x=variable, y=value))+geom_boxplot(aes(fill=activity))+theme_minimal()+ylim(0,9)

# relation between difficulty, preparedness and satisfaction (flow constructs)
ggplot(fb1_data, aes(x=difficulty, y=satisfaction, col=activity))+geom_jitter(size=4, alpha=0.3)+geom_smooth(se=F)
ggpairs(fb1_data[,6:9],
        lower = list(continuous = wrap("points", alpha = 0.1)))+theme_minimal()

```
```{r fb-plots2}

table(fb2_data$group)
ggplot(fb2_data, aes(x=overall.satisfaction))+geom_density()+theme_minimal()+xlim(0,9)

```

## Draw some plots - group observations

Plot group variables over time...

```{r group-obs, echo=FALSE}

# ggplot(accdata, aes(x=timestamp, colour="x")) + 
#   geom_line(aes(y = x, colour = "x"))+ 
#   geom_line(aes(y = y, colour = "y"))+ 
#   geom_line(aes(y = z, colour = "z"))+theme_minimal()+
#   ggtitle("Accelerometer X, Y, Z over time")

#melted <- melt(group_obs, id=1:2, measure=4:7)

ggplot(group_obs, aes(x=timestamp, y=talking, col=as.factor(group)))+geom_smooth(se = F)+theme_minimal()


```

Let's do PCA


```{r pca}

pca.group=PCA(group_obs[,4:8], scale.unit = TRUE, graph = TRUE)


group_obs$engagement <- pca.group$ind$coord[,1]
group_obs$focusint <- pca.group$ind$coord[,2]

pca.group=PCA(group_obs[,4:8], scale.unit = TRUE, graph = TRUE)


```


```{r pca2}


fviz_eig(pca.group, addlabels = TRUE)

var <- get_pca_var(pca.group)

fviz_pca_ind(pca.group,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = group_obs$group, # color by groups
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups", ellipse.type = "confidence", repel = T
             )

```



## Draw some plots - student 

Plot group variables over time...

```{r student-obs, echo=FALSE}

# ggplot(accdata, aes(x=timestamp, colour="x")) + 
#   geom_line(aes(y = x, colour = "x"))+ 
#   geom_line(aes(y = y, colour = "y"))+ 
#   geom_line(aes(y = z, colour = "z"))+theme_minimal()+
#   ggtitle("Accelerometer X, Y, Z over time")

#melted <- melt(group_obs, id=1:2, measure=4:7)
student_obs$student.id <- paste(student_obs$group,student_obs$student)
ggplot(student_obs, aes(x=timestamp, y=talking, col=as.factor(student.id)))+geom_smooth(se = F)+theme_minimal()


```

Let's do PCA


```{r pca.stu}

pca.student=PCA(student_obs[,4:8], scale.unit = TRUE, graph = TRUE)

# Interpretation of dimensions: engagement, external focus
student_obs$engagement <- pca.student$ind$coord[,1]
student_obs$focusext <- pca.student$ind$coord[,2]

pca.student=PCA(student_obs[,4:8], scale.unit = TRUE, graph = TRUE)

```


```{r pca2.stu}


fviz_eig(pca.student, addlabels = TRUE)

var <- get_pca_var(pca.student)

print(var$cor)

fviz_pca_ind(pca.student,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = student_obs$student.id, # color by groups
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Students", ellipse.type = "confidence", repel = T
             )

ag.stu <- aggregate(cbind(engagement,focusext)~student.id, data=student_obs, FUN=mean)
ag.stu

ggplot(student_obs, aes(x=engagement, col=group))+geom_density()+theme_minimal()

ggplot(student_obs, aes(x=timestamp, y=engagement))+geom_smooth(span = 0.2)+theme_minimal()#+geom_point(alpha=0.1,size=4)

ggplot(student_obs, aes(x=timestamp, y=focusext))+geom_smooth(span = 0.2)+theme_minimal()#+geom_point(alpha=0.1,size=4)


```
























Total movement over time (calculated as the difference in all 3 axes to the previous value)

```{r movement, echo=FALSE}
totalacc <- data.frame()
for(group in unique(acceleration$group)){
  accdata <- acceleration[acceleration$group==group,]
  accdata$diff <- sqrt((accdata$x-shift(accdata$x))^2+
                   (accdata$y-shift(accdata$y))^2+
                   (accdata$z-shift(accdata$z))^2)
  if(nrow(totalacc)==0) totalacc <- accdata
  else totalacc <- rbind(totalacc, accdata)
}



ggplot(totalacc, aes(x=timestamp, y=diff, col=as.factor(group)))+geom_smooth(se=F)+theme_minimal()+
  ggtitle("Total movement over time")


```

Distance (aprox.) to the beacons over time

```{r beacons, echo=FALSE}

ggplot(beacons, aes(x=timestamp, y=rssi, colour=minor))+
   geom_line()+theme_minimal()+facet_wrap(~as.factor(group))
   ggtitle("Beacon signal strength")


```

And some geoposition data... (no useful data yet)

```{r geo, echo=FALSE}

# #ggplot(geodata, aes(x=timestamp, y=latitude))+
# #  geom_line(aes(y = latitude, colour = "latitude"))+ 
# #  geom_line(aes(y = longitude, colour = "longitude"))+ 
# #  theme_minimal()+
# #  ggtitle("Lat/Long over time")
# 
# ggplot(geodata, aes(x=longitude, y=latitude))+
#   geom_line()+ 
#   theme_minimal()+
#   ggtitle("Lat/Long over time")


```
